04_analysis

### 2023-11-21 05_analysis.Rmd
Update 05_analysis.Rmd

@@

# Which buoy was the loudest?
```{r loudest-buoy}
allDrifts %>%
  ggplot(mapping = aes(x = Deployment, y = TOL_2000)) + geom_boxplot() + theme_bw()

@@

#Plot facet grid
ggplot(ZC_pivot, aes(x = UTC, y = value, color = avg)) + geom_smooth() +
  facet_grid(vars(avg), scales = "free") +
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Hourly Averages of ZC Detections and Soundscape Metrics -  
       Buoy 21")
```
# rollapply for vessels
```{r}
ZC_filt_ship <- GPSwhaleAISenvSS08 %>% 
  filter(species %in% c("ZC", NA)) %>%
  # Hourly rolling averages
  mutate(rollAvg_Whl = zoo::rollapply(Wpresence, 5, mean, fill = NA)) %>%
  mutate(rollAvg_2000 = zoo::rollapply(TOL_2000, 5, mean, fill = NA)) %>%
  mutate(rollAvg_125 = zoo::rollapply(TOL_125, 5, mean, fill = NA)) %>%
  mutate(rollAvg_ship = zoo::rollapply(shipin10km, 5, mean, fill = NA)) %>%
  # Drop first and last NAs that are created as part of averages
  drop_na(., rollAvg_Whl)

@@

#Plot facet grid
ggplot(ZC_pivot_ship, aes(x = UTC, y = value, color = avg)) + geom_smooth() +
  facet_grid(vars(avg), scales = "free") +
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Hourly Averages of ZC Detections + SS Metrics + Ships -  
       Buoy 08")
```

Vessel presence vs ZC presence
```{r}
ZC_filt_ship <- GPSwhaleAISenvSS23 %>% 
  filter(species %in% c("ZC", NA)) %>%
  # Hourly rolling averages
  mutate(rollAvg_Whl = zoo::rollapply(Wpresence, 5, mean, fill = NA)) %>%
  mutate(rollAvg_2000 = zoo::rollapply(TOL_2000, 5, mean, fill = NA)) %>%
  mutate(rollAvg_125 = zoo::rollapply(TOL_125, 5, mean, fill = NA)) %>%
  mutate(rollAvg_ship = zoo::rollapply(shipin10km, 5, mean, fill = NA)) %>%
  # Drop first and last NAs that are created as part of averages
  drop_na(., rollAvg_Whl)

# One plot
ggplot(ZC_filt_ship, aes(x = UTC)) +
  geom_smooth(aes(y = rollAvg_ship, color = "ship presence")) +
  geom_smooth(aes(y = rollAvg_Whl, color = "whale")) +  
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Hourly Averages of ZC Detections and Ships -  
       Buoy 23")

@@

# rollApply on same plot for ZC
```{r}
# Plot lines on same plot

min_first <- min(ZC_filt$rollAvg_Whl)
max_first <- max(ZC_filt$rollAvg_Whl)
min_second <- min(ZC_filt$rollAvg_2000)
max_second <- max(ZC_filt$rollAvg_2000)

@@

# Plot two lines on same plot
ggplot(ZC_filt, aes(x = UTC)) +
  geom_smooth(aes(y = rollAvg_Whl, color = "ZC presence")) +
  geom_smooth(aes(y = inv_scale_function(rollAvg_2000, scale, shift), color = "TOL 2000")) +
  scale_x_datetime(date_breaks = "3 days") +
  scale_y_continuous(limits = c(min_first, max_first), 
                     sec.axis = sec_axis(~scale_function(., scale, shift), 
                                         name = "TOL 2000 - hourly rolling average")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Averages of ZC Detections alongside Soundscape Metrics - Buoy 13")

########################################
## Manual scale
scale = 2000
ggplot(ZC_filt, aes(x = UTC)) +
  geom_smooth(aes(y = rollAvg_Whl, color = "ZC presence")) +
  geom_smooth(aes(y = rollAvg_2000/scale, color = "TOL 2000")) +
  scale_x_datetime(date_breaks = "3 days") +
  scale_y_continuous(#limits = c(0.0, 0.05),
                     sec.axis = sec_axis(~.* scale, name = "TOL 2000 - hourly rolling average")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Averages of ZC Detections alongside Soundscape Metrics - Buoy 13")

@@

#Sound levels for all species
```{r}
# Sound levels for all species
ggplot(detsOnly3) + geom_boxplot(aes(x = species, y = TOL_125)) +
  labs(title = "TOL 125 Sound levels by Species - All Buoys") +
  theme_classic()
ggplot(detsOnly3) + geom_boxplot(aes(x = species, y = TOL_2000)) +
    labs(title = "TOL 2000 Sound levels by Species - All Buoys") +
  theme_classic()

>>>>

# Remove points where buoy is on the shelf - 300 meters
allDrifts <- allDrifts %>%
  filter(depth > 300)

@@

# Which buoy was the loudest?
```{r loudest-buoy}
allDrifts %>%
  ggplot(mapping = aes(x = Deployment, y = TOL_2000)) + 
  geom_violin() + 
  geom_boxplot(width = 0.1) +
  theme_bw()

@@

#Plot facet grid
ggplot(ZC_pivot, aes(x = UTC, y = value, color = avg)) + geom_line() +
  facet_grid(vars(avg), scales = "free") +
  scale_x_datetime(date_breaks = "1 days") +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5, hjust=1),
        legend.position = "none") +
  labs(title = "Rolling Hourly Averages of ZC Detections and   
  Soundscape Metrics - Buoy 21")
```
# rollapply for vessels
```{r}
ZC_filt_ship <- GPSwhaleAISSS20.alt%>% 
  filter(species %in% c("ZC", NA)) %>%
  # Hourly rolling averages
  mutate(rollAvg_Whl = zoo::rollapply(Wpresence, 5, mean, fill = NA)) %>%
  mutate(rollAvg_2000 = zoo::rollapply(TOL_2000, 5, mean, fill = NA)) %>%
  mutate(rollAvg_125 = zoo::rollapply(TOL_125, 5, mean, fill = NA)) %>%
  mutate(rollAvg_ship = zoo::rollapply(shipin10km, 5, mean, fill = NA)) %>%
  # Drop first and last NAs that are created as part of averages
  drop_na(., rollAvg_Whl)

@@

#Plot facet grid
ggplot(ZC_pivot_ship, aes(x = UTC, y = value, color = avg)) + geom_smooth() +
  facet_grid(vars(avg), scales = "free") +
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(#angle = 90, 
                                   vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Hourly Averages of ZC Detections + SS Metrics + Ships -  
       Buoy 20")
```

Original data points plotted
```{r}
ZC_pivot_reg <- pivot_longer(ZC_filt, cols = c(TOL_125,TOL_2000, Wpresence), names_to = "datacol", values_to = "value")

# Regular plotted detections
ggplot(ZC_pivot_reg, aes(x = UTC, y = value, color = datacol)) + geom_point(size = 0.1) +
  facet_grid(vars(datacol), scales = "free") +
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5, hjust=1)) +
  labs(title = "ZC Detections + SS Metrics - Buoy 21")

ggplot(ZC_pivot_reg, aes(x = UTC, y = value, color = datacol)) + geom_line() +
  facet_grid(vars(datacol), scales = "free") +
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90,
                                   vjust = 0.5, hjust=1)) +
  labs(title = "ZC Detections + SS Metrics + Ships - Buoy 13")

@@

Vessel presence vs ZC presence
```{r}
ZC_filt_ship <- GPSwhaleAISSS20.alt %>% 
  filter(species %in% c("ZC", NA)) %>%
  # Hourly rolling averages
  mutate(rollAvg_Whl = zoo::rollapply(Wpresence, 5, mean, fill = NA)) %>%
  mutate(rollAvg_2000 = zoo::rollapply(TOL_2000, 5, mean, fill = NA)) %>%
  mutate(rollAvg_125 = zoo::rollapply(TOL_125, 5, mean, fill = NA)) %>%
  mutate(rollAvg_ship = zoo::rollapply(shipin10km, 5, mean, fill = NA)) %>%
  # Drop first and last NAs that are created as part of averages
  drop_na(., rollAvg_Whl)

# One plot
ggplot(ZC_filt_ship, aes(x = UTC)) +
  geom_smooth(aes(y = rollAvg_ship, color = "ship presence")) +
  geom_smooth(aes(y = rollAvg_Whl, color = "whale")) +  
  scale_x_datetime(date_breaks = "3 days") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Hourly Averages of ZC Detections and Ships -  
       Buoy 20")

# rollApply on same plot for ZC
```{r}
# Read in buoy 20 without environmental variables
GPSwhaleAISSS20.alt <- readRDS("D:/Buoy_dfs/GPSwhaleAISSS20_alt.rda")

ZC_filt_a <- GPSwhaleAISSS20.alt%>% 
  filter(species %in% c("ZC", NA)) %>%
  # Hourly rolling averages
  mutate(rollAvg_Whl = zoo::rollapply(Wpresence, 5, mean, fill = NA)) %>%
  mutate(rollAvg_2000 = zoo::rollapply(TOL_2000, 5, mean, fill = NA)) %>%
  mutate(rollAvg_125 = zoo::rollapply(TOL_125, 5, mean, fill = NA)) %>%
  # Drop first and last NAs that are created as part of averages
  drop_na(., rollAvg_Whl)

# Plot lines on same plot
min_first <- min(ZC_filt_a$rollAvg_Whl)
max_first <- max(ZC_filt_a$rollAvg_Whl)
min_second <- min(ZC_filt_a$rollAvg_2000)
max_second <- max(ZC_filt_a$rollAvg_2000)

@@

# Plot two lines on same plot
ggplot(ZC_filt_a, aes(x = UTC)) +
  geom_line(aes(y = rollAvg_Whl, color = "ZC presence")) +
  geom_line(aes(y = inv_scale_function(rollAvg_2000, scale, shift), color = "TOL 2000")) +
  scale_x_datetime(date_breaks = "3 days") +
  scale_y_continuous(limits = c(min_first, max_first), 
                     sec.axis = sec_axis(~scale_function(., scale, shift), 
                                         name = "TOL 2000 - hourly rolling average")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Averages of ZC Detections alongside   
       Soundscape Metrics - Buoy 20")
```

```{r}
########################################
## Manual scale
scale = 2000
ggplot(ZC_filt, aes(x = UTC)) +
  geom_smooth(aes(y = rollAvg_Whl, color = "ZC presence")) +
  geom_smooth(aes(y = rollAvg_2000/scale, color = "TOL 2000")) +
  scale_x_datetime(date_breaks = "3 days") +
  scale_y_continuous(#limits = c(0.0, 0.05),
                     sec.axis = sec_axis(~.* scale, name = "TOL 2000 - hourly rolling average")) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "Rolling Averages of ZC Detections alongside Soundscape Metrics - Buoy 13")

@@

#Sound levels for all species
```{r}
# Sound levels for all species
ggplot(detsOnly3) + geom_violin(aes(x = species, y = TOL_125)) +
  labs(title = "TOL 125 Sound levels by Species - All Buoys") +
  theme_classic()

ggplot(detsOnly3, aes(x = species, y = TOL_2000)) + geom_violin() +
  labs(title = "TOL 2000 Sound levels by Species - All Buoys") +
  theme_classic() +
  geom_boxplot(width = 0.1)
